---
title: "Simulation Exercise"
author: "Sam Stites"
date: "November 22, 2014"
output: pdf_document
---
### Question
The exponential distribution can be simulated in R with `rexp(n, lambda)` where lambda is the rate parameter. The mean of exponential distribution is `1/lambda` and the standard deviation is also also `1/lambda`. Set `lambda = 0.2` for all of the simulations. In this simulation, you will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand or so simulated averages of 40 exponentials.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials.  You should

1. Show where the distribution is centered at and compare it to the theoretical center of the distribution.
2. Show how variable it is and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.

Note that for point 3, focus on the difference between the distribution of a large collection of random exponentials and the distribution of a large collection of averages of 40 exponentials.

### Part 1

Given the following criteria:
```{r}
set.seed(7)
n             <-   40
lambda        <-  0.2
nsimultations <- 1000
simulation    <- replicate(n = nsimultations,
                           expr = rexp(n,lambda))
```

We see that there the simulated distribution of means is centered around 4.983294:

```{r}
mean(simulation)
#=> [1] 4.983294
```

Very close to the exponential distribution's theoretical mean, `1/lamdba`, for a lambda of `0.2`:

```{r}
1/lambda
#=> [1] 5
```

### Part 2

Similarly, the standard deviation of the theoretical distribution is also `1/lambda`, `5` for a lambda of 0.2, which is also quite close to the simulated standard of deviatiation:

```{r}
sd(simulation)
#=> [1] 4.984027
```

### Part 3
Given the dataset, we can see that the distribution of a large collection of random exponentials takes on an exponential distribution and that the distribution of the large collection of averages of 40 exponentials takes on an approximate normal distribution — as would be expected by the central limit theorem. Below are plots of the data and their respective densities - to see the generating functions, look at the appendix.

``` {r, echo=FALSE, fig.width = 3, fig.height = 3}
means <- colMeans(simulation)
makeHist<-function(data, title){
  hist(data,
       freq=FALSE,
       col = "light blue",
       breaks=20,
       main=title)
  lines(density(data), col = "red")
}
makeHist(means,
"Dist. of large
collection of means for
40 exponentials")
makeHist(simulation,
"Dist. of large
collection of random
exponentials")
```

Taking a look at quantile-to-quantile comparisons, we see that our expectations are accurate:
``` {r, echo=FALSE, fig.width = 3, fig.height = 3}
qqnorm(means,
       col = "blue",
ylab="Mean Quantiles",
main="normal QQ
with means");
qqline(means, col = "red")
# compare simulations to the exponential distributions
qqplot(rexp(1000, rate=lambda),
       simulation,
       col = "blue",
main="exponential QQ
with simulations",
xlab="Theoretical Quantiles",
ylab="Simulation Quantiles")
abline(0,1, col = "red")
```

### Appendix

``` {r}
# Part 3
means <- colMeans(simulation)
makeHist<-function(data, title){
  hist(data,
       freq=FALSE,
       col = "light blue",
       breaks=20,
       main=title)
  lines(density(data), col = "red")
}
makeHist(means,      "Dist. of large collection of means for 40 exponentials")
makeHist(simulation, "Dist. of a large collection of random exponentials")
# Q-Q plots
# compare means to normal distribution
qqnorm(means,
       col = "blue",
       main="theoretical normal QQ");
qqline(means, col = "red")
# compare simulations to the exponential distributions
qqplot(rexp(1000, rate=lambda),
       simulation,
       col = "blue",
       main="theoretical exponential QQ")
abline(0,1, col = "red")
```
